---
title: "Rotated LC Optimization"
output: html_document
---

# 1. Preparing Data

Sourcing external custom functions:

```{r}
# In case we need the parent directorcy
#parts <- unlist(strsplit(getwd(), .Platform$file.sep))
#ParentDirectory <- do.call(file.path, as.list(parts[1:length(parts) - 1]))

source("RotatedLC_Functions.R")
```

Loading and preparing HMD data:


```{r}

country <- "ISL"

DataOutput <- CleanUp_HMD(country, 1950, 1990, 1999)
```
Check the output m_x tables

```{r}

nrow(DataOutput$mx_male_train)
nrow(DataOutput$mx_male_validation)
nrow(DataOutput$mx_male_test)
colnames(DataOutput$mx_male_train)
colnames(DataOutput$mx_male_validation)
colnames(DataOutput$mx_male_test)

nrow(DataOutput$mx_female_train)
nrow(DataOutput$mx_female_validation)
nrow(DataOutput$mx_female_test)
colnames(DataOutput$mx_female_train)
colnames(DataOutput$mx_female_validation)
colnames(DataOutput$mx_female_test)

```

# 2. Rotated Lee-Carter

Checking the $e_0$ projectons of the plain L-C model:

```{r}
testresults <- ForecastRotatedLC(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                e0_female_proj = NULL, e0_male_proj = NULL,
                                0.5, 87, ncol(DataOutput$mx_male_validation)+ncol(DataOutput$mx_male_test), IsLog = TRUE)

testresults$est_e0_male
testresults$est_e0_female

(testresults$est_e0_male+testresults$est_e0_female)/2

testresults$arima_male_coef
testresults$arima_female_coef
```


Executing optimisation of $p$ on HMD.
**!!!IMPORTANT!!!** Parameters e0_female_proj and e0_male_proj must be **NULL**!!

```{r}
OptimisationResults <- OptimizeRotatedLC(mx_male_train = DataOutput$mx_male_train,
                                         mx_female_train = DataOutput$mx_female_train,
                                         mx_male_validation = DataOutput$mx_male_validation,
                                         mx_female_validation = DataOutput$mx_female_validation,
                                         mx_male_test = DataOutput$mx_male_test,
                                         mx_female_test = DataOutput$mx_female_test,
                                         nx = 1,
                                         e0_female_proj = NULL,
                                         e0_male_proj = NULL,
                                         is_log = TRUE)


OptimisationResults$optimal_e0l_male
OptimisationResults$optimal_p_male

OptimisationResults$optimal_e0l_female
OptimisationResults$optimal_p_female
```

Plotting the RMSE values along $p$ and $e_0$ for both sexes:

```{r}
mse_results <- OptimisationResults$mse_table

library(ggplot2)

ggplot(mse_results, aes(e0l, p, fill= mse_female_validation)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of p and e0l on Validation Set - Females")

ggplot(mse_results, aes(e0l, p, fill= mse_male_validation)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of p and e0l on Validation Set - Males")


ggplot(mse_results, aes(e0l, p, fill= mse_female_test)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of p and e0l on Test Set - Females")

ggplot(mse_results, aes(e0l, p, fill= mse_male_test)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of p and e0l on Test Set - Males")



```



Finding optimal $p$ for all countries based on HMD Data:

```{r}
CountryList <- c("AUS", "AUT", "BLR", "BEL", "BGR", "CAN", "CZE", "DNK", "EST", "FIN", "FRATNP", "DEUTE", "DEUTW", "GRC", "HKG", "HUN", "ISL", "IRL", "ISR", "ITA", "JPN", "LVA", "LTU", "LUX", "NLD", "NZL_NP", "NOR", "POL", "PRT", "RUS", "SVK", "SVN", "ESP", "SWE", "CHE", "TWN", "GBR_NP", "GBRTENW", "GBR_SCO", "GBR_NIR", "USA", "UKR")

ResultList <- vector(mode = "list", length = length(CountryList))
names(ResultList) <- CountryList

for (index in 1:length(CountryList)) {
  DataOutput <- CleanUp_HMD(CountryList[index], 1950, 1990, 1999)
  
  print(CountryList[index])
  
  tryCatch(
    expr = {
      OptimisationResults <- OptimizeRotatedLC(mx_male_train = DataOutput$mx_male_train,
                                         mx_female_train = DataOutput$mx_female_train,
                                         mx_male_validation = DataOutput$mx_male_validation,
                                         mx_female_validation = DataOutput$mx_female_validation,
                                         mx_male_test = DataOutput$mx_male_test,
                                         mx_female_test = DataOutput$mx_female_test,
                                         nx = 1,
                                         e0_female_proj = NULL,
                                         e0_male_proj = NULL,
                                         is_log = TRUE)
      ResultList[[index]] <- OptimisationResults
    },
    error = function(error_condition){ 
        ResultList[[index]] <- error_condition
    }
  )
}
```


Save results to RDA:

```{r}
save(ResultList,file="HMD_RotatedLC_Optimised_p_e0l_LogMSE.Rda")
```


Plotting the RMSE values along $p$ and $e_0$ for both sexes:

```{r}
OptimisationResults <- ResultList$SWE

mse_results <- OptimisationResults$mse_table

library(ggplot2)

ggplot(mse_results, aes(e0l, p, fill= mse_female_validation)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of p and e0l on Validation Set - Females")

ggplot(mse_results, aes(e0l, p, fill= mse_male_validation)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of p and e0l on Validation Set - Males")


ggplot(mse_results, aes(e0l, p, fill= mse_female_test)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of p and e0l on Test Set - Females")

ggplot(mse_results, aes(e0l, p, fill= mse_male_test)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of p and e0l on Test Set - Males")

```

# 3. Lee-Carter with GAM

Testing the function for LC+GAM $m_x$ forecasts with fixed parameters

```{r}
testresults <- ForecastLCwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = 7, basis_form = "ts",
                                 ForecastPeriods = ncol(DataOutput$mx_male_validation)+ncol(DataOutput$mx_male_test),
                                 IsLog = TRUE)

testresults$arima_male_coef
testresults$arima_female_coef

plotmo::plotmo(testresults$GAM_male)
plotmo::plotmo(testresults$GAM_female)
```

Executing optimisation of $k$ and $bs$ on HMD.

```{r}
OptimisationResults <- OptimizeLCwithGAM(mx_male_train = DataOutput$mx_male_train,
                                         mx_female_train = DataOutput$mx_female_train,
                                         mx_male_validation = DataOutput$mx_male_validation,
                                         mx_female_validation = DataOutput$mx_female_validation,
                                         mx_male_test = DataOutput$mx_male_test,
                                         mx_female_test = DataOutput$mx_female_test,
                                         nx = 1,
                                         is_log = TRUE)
```

Plotting the RMSE values along $k$ and $bs$ for both sexes:

```{r}
mse_results <- OptimisationResults$mse_table

library(ggplot2)

ggplot(mse_results, aes(x=basis_form,y=k_base, fill= mse_female_validation)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of k and bs on Validation Set - Females")

ggplot(mse_results, aes(basis_form, k_base, fill= mse_male_validation)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of k and bs on Validation Set - Males")


ggplot(mse_results, aes(basis_form, k_base, fill= mse_female_test)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of k and bs on Test Set - Females")

ggplot(mse_results, aes(basis_form, k_base, fill= mse_male_test)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of k and bs on Test Set - Males")



```

Finding optimal $k$ and $bs$ for all countries based on HMD Data:

```{r}
#CountryList <- c("AUS", "AUT", "BLR", "BEL", "BGR", "CAN", "CZE", "DNK", "EST", "FIN", "FRATNP", "DEUTE", "DEUTW", "GRC", #"HKG", "HUN", "ISL", "IRL", "ISR", "ITA", "JPN", "LVA", "LTU", "LUX", "NLD", "NZL_NP", "NOR", "POL", "PRT", "RUS", "SVK", #"SVN", "ESP", "SWE", "CHE", "TWN", "GBR_NP", "GBRTENW", "GBR_SCO", "GBR_NIR", "USA", "UKR")

CountryList <- c("BLR", "EST", "DEUTE", "DEUTW", "GRC", "HKG", "ISL", "TWN", "GBRTENW")

ResultList <- vector(mode = "list", length = length(CountryList))
names(ResultList) <- CountryList

for (index in 1:length(CountryList)) {
  DataOutput <- CleanUp_HMD(CountryList[index], 1950, 1990, 1999)
  
  print(CountryList[index])
  
  tryCatch(
    expr = {
      OptimisationResults <- OptimizeLCwithGAM(mx_male_train = DataOutput$mx_male_train,
                                         mx_female_train = DataOutput$mx_female_train,
                                         mx_male_validation = DataOutput$mx_male_validation,
                                         mx_female_validation = DataOutput$mx_female_validation,
                                         mx_male_test = DataOutput$mx_male_test,
                                         mx_female_test = DataOutput$mx_female_test,
                                         nx = 1,
                                         is_log = TRUE)
      ResultList[[index]] <- OptimisationResults
    },
    error = function(error_condition){ 
        ResultList[[index]] <- error_condition
    }
  )
}

save(ResultList,file="HMD_LCwithGAM_Optimised_k_bs_LogMSE.Rda")
```

Load the exported result list from `Rda` file:
```{r}
load("HMD_LCwithGAM_Optimised_k_bs_LogMSE.Rda")
```


Plotting the RMSE values along $k$ and $bs$ for both sexes:

```{r}
OptimisationResults <- ResultList$GRC

mse_results <- OptimisationResults$mse_table

library(ggplot2)

ggplot(mse_results, aes(x=basis_form,y=k_base, fill= mse_female_validation)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of k and bs on Validation Set - Females")

ggplot(mse_results, aes(basis_form, k_base, fill= mse_male_validation)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of k and bs on Validation Set - Males")


ggplot(mse_results, aes(basis_form, k_base, fill= mse_female_test)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of k and bs on Test Set - Females")

ggplot(mse_results, aes(basis_form, k_base, fill= mse_male_test)) + 
  geom_tile() +
  scale_fill_gradient(low="green", high="red") +
  labs(title = "MSE as a function of k and bs on Test Set - Males")

```

Load the the best results per country from `Rda` file:
```{r}
load("HMD_BestResultsByCountry_Rotated+GAM.Rda")
```

Plot the differences in Male+Female MSE on test set for each Country:

```{r}
library(ggplot2)

HMD_BestResultsByCountry_ToPlotMale <- HMD_BestResultsByCountry[,c("Country", "mse_male_test_RotatedLC", "mse_male_test_LCwithGAM")]
HMD_BestResultsByCountry_ToPlotMale$mse_male_test_Difference <- HMD_BestResultsByCountry_ToPlotMale$mse_male_test_LCwithGAM -
  HMD_BestResultsByCountry_ToPlotMale$mse_male_test_RotatedLC
HMD_BestResultsByCountry_ToPlotMale$WhichIsBetter <- ifelse(HMD_BestResultsByCountry_ToPlotMale$mse_male_test_Difference > 0,
                                                            "RotatedLC", "LCwithGAM")

ggplot(HMD_BestResultsByCountry_ToPlotMale, aes(x=reorder(Country,mse_male_test_Difference), y=mse_male_test_Difference, fill=WhichIsBetter)) +
  geom_bar(stat='identity', position='dodge') +
  labs(title = "Difference in MSE of RotatedLC and LC with GAM on Test Set - Males",
       x=NULL) + coord_flip()

HMD_BestResultsByCountry_ToPlotFemale <- HMD_BestResultsByCountry[,c("Country", "mse_female_test_RotatedLC", "mse_female_test_LCwithGAM")]
HMD_BestResultsByCountry_ToPlotFemale$mse_female_test_Difference <- HMD_BestResultsByCountry_ToPlotFemale$mse_female_test_LCwithGAM -
  HMD_BestResultsByCountry_ToPlotFemale$mse_female_test_RotatedLC
HMD_BestResultsByCountry_ToPlotFemale$WhichIsBetter <- ifelse(HMD_BestResultsByCountry_ToPlotFemale$mse_female_test_Difference > 0,"RotatedLC", "LCwithGAM")

ggplot(HMD_BestResultsByCountry_ToPlotFemale, aes(x=reorder(Country,mse_female_test_Difference), y=mse_female_test_Difference, fill=WhichIsBetter)) +
  geom_bar(stat='identity', position='dodge') +
  labs(title = "Difference in MSE of RotatedLC and LC with GAM on Test Set - Females",
       x=NULL) + coord_flip()

```