---
title: "Spatial Analysis of Best Models"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# 1. Load Data on Model Results for all Country

Loading the *final_results.csv* to R enivronment:

```{r}
HMD_BestResults <- read.csv("final_results.csv")
str(HMD_BestResults)
```

Rename the `variable` column to `Model` and the `mse` to `MSE`. Rename *preds* to *NN*, *preds_boost* to *NN_Boost*, *stacked* to *Stacked* and *pred_LC_svd* to *LC_SVD* to have more clear names on the graphs and pivots.

```{r}
colnames(HMD_BestResults)[3] <- "Model"
colnames(HMD_BestResults)[4] <- "MSE"
HMD_BestResults$Model[HMD_BestResults$Model == "preds"] <- "NN"
HMD_BestResults$Model[HMD_BestResults$Model == "preds_boost"] <- "NN_Boost"
HMD_BestResults$Model[HMD_BestResults$Model == "pred_LC_svd"] <- "LC_SVD"
HMD_BestResults$Model[HMD_BestResults$Model == "stacked"] <- "Stacked"
str(HMD_BestResults)
```

Load packages for analysis:

```{r}
library(ggplot2)
library(tidyr)
library(stringr)
library(dplyr)
```

Define function to shape the `HMD_BestResults` to a *one country - one row* styled table.

```{r}
myspread <- function(df, key, value) {
  # quote key
  keyq <- rlang::enquo(key)
  # break value vector into quotes
  valueq <- rlang::enquo(value)
  s <- rlang::quos(!!valueq)
  df %>% gather(variable, value, !!!s) %>%
    unite(temp, !!keyq, variable) %>%
    spread(temp, value)
}
```

Apply the function on `HMD_BestResults` and see the results:

```{r}
HMD_BestResults <- HMD_BestResults %>% spread(Sex, MSE)
HMD_BestResults <- HMD_BestResults %>% myspread(Model, c(Female,Male))
HMD_BestResults <- HMD_BestResults[,c("Country",
                                      colnames(HMD_BestResults)[str_detect(colnames(HMD_BestResults), "_Male")],
                                      colnames(HMD_BestResults)[str_detect(colnames(HMD_BestResults), "_Female")])]
str(HMD_BestResults)
```

# 2. Get Best Models and Generate New Variables

Get best model for females:

```{r}
HMD_BestResults$BestFemaleMSE <- apply(HMD_BestResults[,8:13], 1, FUN = min)


HMD_BestResults$BestFemaleModel <- NA
for (i in 1:nrow(HMD_BestResults)) {
  HMD_BestResults$BestFemaleModel[i] <- names(HMD_BestResults[,2:13])[which(
    HMD_BestResults[,2:13] == HMD_BestResults$BestFemaleMSE[i],
    arr.ind=T)[, "col"]]
}

str(HMD_BestResults)
```

Get best model for males:

```{r}
HMD_BestResults$BestMaleMSE <- apply(HMD_BestResults[,2:7], 1, FUN = min)


HMD_BestResults$BestMaleModel <- NA
for (i in 1:nrow(HMD_BestResults)) {
  HMD_BestResults$BestMaleModel[i] <- names(HMD_BestResults[,1:9])[which(
    HMD_BestResults[,1:9] == HMD_BestResults$BestMaleMSE[i],
    arr.ind=T)[, "col"]]
}

str(HMD_BestResults)
```

See the best model frequencies:

```{r}
table(HMD_BestResults$BestFemaleModel)
table(HMD_BestResults$BestMaleModel)
```

Looks like *NN_Boost* and *Stacked* are the best option for *females*. And the *LCwithGAM* for *males*.

Let's check the *NN_Boost* advantage to the *simple NN* solution in each country for both genders:

```{r}
temp_df <- reshape2::melt(HMD_BestResults[,c("Country", "NN_Male", "NN_Boost_Male")],
                            id.var = "Country")

ggplot(temp_df, aes(x = Country, y = value, color = variable, group=variable)) +
  geom_line()

temp_df <- reshape2::melt(HMD_BestResults[,c("Country", "NN_Female", "NN_Boost_Female")],
                            id.var = "Country")

ggplot(temp_df, aes(x = Country, y = value, color = variable, group=variable)) +
  geom_line()
```

It seems that the $MSE$ of *NN_Boost* is always *better or about the same* as the $MSE$ of the *simple NN*.

So, next we generate a variable that measures the *NN_boost* advantage in $MSE$ compared to the best **not NN** model.

```{r}
HMD_BestResults$BestMaleMSE_notNN <- apply(HMD_BestResults[,c(2,3,6,7)], 1, FUN = min)
HMD_BestResults$NN_boost_vs_Best_notNN_male <- HMD_BestResults$NN_Boost_Male - HMD_BestResults$BestMaleMSE_notNN

HMD_BestResults$BestFemaleMSE_notNN <- apply(HMD_BestResults[,c(8,9,12,13)], 1, FUN = min)
HMD_BestResults$NN_boost_vs_Best_notNN_female <- HMD_BestResults$NN_Boost_Female - HMD_BestResults$BestFemaleMSE_notNN

str(HMD_BestResults)
```

# 3. Draw Some Maps

See the best models in each country on the world map!

First, download the shapefile:

```{r eval=FALSE}
# Download the shapefile.
download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="GeoData/world_shape_file.zip")

# Unzip the file.
system("unzip GeoData/world_shape_file.zip")
```

Load the shapefile to R:

```{r}
library(rgdal)

# Rread the shape file
world_spdf <- readOGR( 
  dsn= paste0(getwd(),"/GeoData") , 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

# Save the original order of countries so that we don't mess up the shapefile with the merge
world_spdf@data$OriginalOrder <- 1:nrow(world_spdf@data)
```

Cleaning the country codes in the `HMD_BestResults` dataframe:

```{r}
# Creating the dataframe to join
HMD_ResultsToMap <- HMD_BestResults[,c("Country", "BestFemaleModel", "BestMaleModel",
                                       "NN_boost_vs_Best_notNN_male", "NN_boost_vs_Best_notNN_female")]

# Unify some country codes
HMD_ResultsToMap$Country[HMD_ResultsToMap$Country == "FRATNP"] <- "FRA"
HMD_ResultsToMap$Country[HMD_ResultsToMap$Country == "DEUTW"] <- "DEU"
HMD_ResultsToMap$Country[HMD_ResultsToMap$Country == "NZL_NP"] <- "NZL"
HMD_ResultsToMap$Country[HMD_ResultsToMap$Country == "GBR_NP"] <- "GBR"
```

Join with relevant variables from `HMD_ResultsToMap`:

```{r}
ShapeDataTemp <- merge(x=world_spdf@data, y=HMD_ResultsToMap, by.x="ISO3", by.y="Country", all.x=TRUE)

# Clean up the categorical variables and convert them to factor
ShapeDataTemp$BestMaleModel[is.na(ShapeDataTemp$BestMaleModel)] <- "No_Data"
ShapeDataTemp$BestFemaleModel[is.na(ShapeDataTemp$BestFemaleModel)] <- "No_Data"
ShapeDataTemp$BestFemaleModel <- as.factor(ShapeDataTemp$BestFemaleModel)
ShapeDataTemp$BestMaleModel <- as.factor(ShapeDataTemp$BestMaleModel)

world_spdf@data <- ShapeDataTemp[order(ShapeDataTemp$OriginalOrder),]
```

Draw the map of the best models for **female**s!

```{r}
library(leaflet)

# Create a color palette for the map
mypalette <- colorFactor(c("green", "blue", "orange", "grey", "red", "yellow"),
                         world_spdf@data$BestFemaleModel)

# Prepare the text for tooltips
mytext <- paste(
    "Country: ", world_spdf@data$NAME,"<br/>", 
    "Best Model Female: ", world_spdf@data$BestFemaleModel, "<br/>", 
    "NN_boost_vs_Best_notNN_female: ", round(world_spdf@data$NN_boost_vs_Best_notNN_female, 3), 
    sep="") %>%
  lapply(htmltools::HTML)

# Create map object
m <- leaflet(world_spdf) %>% 
  addTiles()  %>% 
  setView( lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(BestFemaleModel), 
    stroke=FALSE,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend(pal=mypalette, values=~BestFemaleModel, opacity=0.9, title = "Best Model for Females", position = "bottomleft")

# Show map object
m

# Save map in a html file
htmlwidgets::saveWidget(m, file="BestFemaleModel.html")
```

Draw the map of the best models for **male**s!

```{r}
library(leaflet)

# Create a color palette for the map
mypalette <- colorFactor(c("green", "blue", "orange", "grey", "red", "yellow"),
                         world_spdf@data$BestMaleModel)

# Prepare the text for tooltips
mytext <- paste(
    "Country: ", world_spdf@data$NAME,"<br/>", 
    "Best Model Male: ", world_spdf@data$BestMaleModel, "<br/>", 
    "NN_boost_vs_Best_notNN_male: ", round(world_spdf@data$NN_boost_vs_Best_notNN_male, 3), 
    sep="") %>%
  lapply(htmltools::HTML)

# Create map object
m <- leaflet(world_spdf) %>% 
  addTiles()  %>% 
  setView( lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(BestMaleModel), 
    stroke=FALSE,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend(pal=mypalette, values=~BestMaleModel, opacity=0.9, title = "Best Model for Males", position = "bottomleft")

# Show map object
m

# Save map in a html file
htmlwidgets::saveWidget(m, file="BestMaleModel.html")
```

Draw the map of *NN_boost* advantage in $MSE$ compared to the *best not NN* model for **female**s!

```{r}
library(leaflet)

# Set the differences on a logaritmic scale
world_spdf@data$Log_NN_boost_vs_Best_notNN_female <- log(world_spdf@data$NN_boost_vs_Best_notNN_female - (min(world_spdf@data$NN_boost_vs_Best_notNN_female, na.rm = TRUE)-0.09))

# Create a color palette for the map
mypalette <- colorNumeric(palette = "RdYlGn",
                          domain = world_spdf@data$Log_NN_boost_vs_Best_notNN_female,
                          na.color="transparent",
                          reverse = TRUE)

# Prepare the text for tooltips
mytext <- paste(
    "Country: ", world_spdf@data$NAME,"<br/>", 
    "Best Model Female: ", world_spdf@data$BestMaleFemodel, "<br/>", 
    "Log_NN_boost_vs_Best_notNN_female: ", round(world_spdf@data$Log_NN_boost_vs_Best_notNN_female, 3), 
    sep="") %>%
  lapply(htmltools::HTML)

# Create map object
m <- leaflet(world_spdf) %>% 
  addTiles()  %>% 
  setView( lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(Log_NN_boost_vs_Best_notNN_female), 
    stroke=TRUE,
    fillOpacity = 0.9, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend(pal=mypalette, values=~Log_NN_boost_vs_Best_notNN_female, opacity=0.9, title = "Log_NN_boost advantage in MSE - Female", position = "bottomleft")

# Show map object
m

# Save map in a html file
htmlwidgets::saveWidget(m, file="NNboostAdvantageFemale.html")
```

Draw the map of *NN_boost* advantage in $MSE$ compared to the *best not NN* model for **male**s!

```{r}
library(leaflet)

# Set the differences on a logaritmic scale
world_spdf@data$Log_NN_boost_vs_Best_notNN_male <- log(world_spdf@data$NN_boost_vs_Best_notNN_male - (min(world_spdf@data$NN_boost_vs_Best_notNN_male, na.rm = TRUE)-0.09))

# Create a color palette for the map
mypalette <- colorNumeric(palette = "RdYlGn",
                          domain = world_spdf@data$Log_NN_boost_vs_Best_notNN_male,
                          na.color="transparent",
                          reverse = TRUE)

# Prepare the text for tooltips
mytext <- paste(
    "Country: ", world_spdf@data$NAME,"<br/>", 
    "Best Model male: ", world_spdf@data$BestMaleFemodel, "<br/>", 
    "Log_NN_boost_vs_Best_notNN_male: ", round(world_spdf@data$Log_NN_boost_vs_Best_notNN_male, 3), 
    sep="") %>%
  lapply(htmltools::HTML)

# Create map object
m <- leaflet(world_spdf) %>% 
  addTiles()  %>% 
  setView( lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(Log_NN_boost_vs_Best_notNN_male), 
    stroke=TRUE,
    fillOpacity = 0.9, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend(pal=mypalette, values=~Log_NN_boost_vs_Best_notNN_male, opacity=0.9, title = "Log_NN_boost advantage in MSE - male", position = "bottomleft")

# Show map object
m

# Save map in a html file
htmlwidgets::saveWidget(m, file="NNboostAdvantageMale.html")
```

# 4. Calculate Moran's I for the NN_Boost Advantage

Load packages for the spatial autocorrelation analysis:

```{r}
library(sf)
library(spdep)
```

Load the shapefile of world countries with the `sf` package:

```{r}
s <- st_read( "GeoData/TM_WORLD_BORDERS_SIMPL-0.3.shp")
```

Add the *NN_boost_vs_Best_notNN_male* and *NN_boost_vs_Best_notNN_female* variables to the `s` shape data.

```{r}
ShapeDataTemp <- merge(x=s, y=HMD_ResultsToMap[,c("Country",
                                                  "NN_boost_vs_Best_notNN_male",
                                                  "NN_boost_vs_Best_notNN_female")],
                       by.x="ISO3", by.y="Country", all.x=TRUE)
s <- ShapeDataTemp
str(s)
```

Ok, we have the two new variables.

Let's define “neighboring” polygons.

```{r}
sf::sf_use_s2(FALSE)
#s <- na.omit(s)
nb <- poly2nb(s, queen=TRUE)
```

Each neighboring polygon will be assigned equal weight when computing the neighboring NN advantages in $MSE$.

```{r}
lw <- nb2listw(nb, style="W", zero.policy=TRUE)
```

Computing the Moran’s $I$ statistic for *Females*.

```{r}
NN_boost_vs_Best_notNN_female_Lag <- lag.listw(lw, s$NN_boost_vs_Best_notNN_female,
                                               zero.policy = FALSE, NAOK = TRUE)

plot(NN_boost_vs_Best_notNN_female_Lag ~ s$NN_boost_vs_Best_notNN_female, pch=16, asp=1)
MoranOLS <- lm(NN_boost_vs_Best_notNN_female_Lag ~ s$NN_boost_vs_Best_notNN_female)
abline(MoranOLS, col="blue")

summary(MoranOLS)
```

Ok, we have some moderate spatial autocorrelation ($0.51$) for *females*. The autpcorrelation is significant at $\alpha=0.05$, but not at $\alpha=0.01$ level.

Computing the Moran’s $I$ statistic for *Males*.

```{r}
NN_boost_vs_Best_notNN_male_Lag <- lag.listw(lw, s$NN_boost_vs_Best_notNN_male,
                                               zero.policy = FALSE, NAOK = TRUE)

plot(NN_boost_vs_Best_notNN_male_Lag ~ s$NN_boost_vs_Best_notNN_male, pch=16, asp=1)
MoranOLS <- lm(NN_boost_vs_Best_notNN_male_Lag ~ s$NN_boost_vs_Best_notNN_male)
abline(MoranOLS, col="blue")

summary(MoranOLS)
```

Ok, we have some moderate spatial autocorrelation ($0.52$) for *males*. The autpcorrelation is significant at $\alpha=0.05$, but not at $\alpha=0.01$ level.

The level of spatial autocorrelation is similar for males and for females, but for males it is seems that some outlier effects are more responsible.