---
title: "Export Mortality Rates"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Read Data on Model Results for all Country

Loading the *csv* to R enivronment:


```{r}
library(readr)
HMD_BestResults <- read_csv("HMD_BestResultsByCountry_Rotated+GAM+NN.csv")
HMD_BestResults <- as.data.frame(HMD_BestResults)
str(HMD_BestResults)
```

Remove columns of valdiation results:

```{r}
HMD_BestResults$min_mse_female_validation_RotatedLC <- NULL
HMD_BestResults$min_mse_male_validation_RotatedLC <- NULL
HMD_BestResults$min_mse_female_validation_LCwithGAM <- NULL
HMD_BestResults$min_mse_male_validation_LCwithGAM <- NULL
str(HMD_BestResults)
```
Load packages for analysis:

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
```

## 2. Get the best RotatedLC parameters for each country.

Load the detailed *RotatedLC* results:

```{r}
load("HMD_RotatedLC_Full_LogMSE.Rda")
str(HMD_RotatedLC_Full)
```

Adjust column names to be compatible with HMD_BestResults:

```{r}
colnames(HMD_RotatedLC_Full)[5:6] <- colnames(HMD_BestResults[2:3])
str(HMD_RotatedLC_Full)
```

Let's get the $e0l$ and $p$ parameters based on the test $MSE$ in the `HMD_BestResults` dataframe:

```{r}
# Males
HMD_Best_withRotatedLC_Male <- HMD_BestResults[,c("Country", "mse_male_test_RotatedLC")]
HMD_Best_withRotatedLC_Male <- merge(HMD_Best_withRotatedLC_Male,
                                     HMD_RotatedLC_Full[,c("Country",
                                                           "mse_male_test_RotatedLC",
                                                           "e0l",
                                                           "p")],
                                     by=c("Country", "mse_male_test_RotatedLC"))



# Aggregate multiple optima
HMD_Best_withRotatedLC_Male <- HMD_Best_withRotatedLC_Male %>% group_by(Country) %>% summarise_all(min)

#----------------------------------------------------------------------------------------------------

# Females
HMD_Best_withRotatedLC_Female <- HMD_BestResults[,c("Country", "mse_female_test_RotatedLC")]
HMD_Best_withRotatedLC_Female <- merge(HMD_Best_withRotatedLC_Female,
                                     HMD_RotatedLC_Full[,c("Country",
                                                           "mse_female_test_RotatedLC",
                                                           "e0l",
                                                           "p")],
                                     by=c("Country", "mse_female_test_RotatedLC"))

# Aggregate multiple optima
HMD_Best_withRotatedLC_Female <- HMD_Best_withRotatedLC_Female %>% group_by(Country) %>% summarise_all(min)

```

## 3. Get the estimated $m_x$ from the best RotatedLC for each country.

Sourcing the necessary functions.

```{r}
source("RotatedLC_MortRateExportFunctions.R")
```

Get the estimated $m_x$ for the training set and forecasted years for one country and one gender.

```{r}
country <- "USA"

DataOutput <- CleanUp_HMD(country, 1950, 1999, 2011)



est_mx_female <- EstimateTrainingRotatedLC(DataOutput$mx_male_train, DataOutput$mx_female_train, 1, DataOutput$e0data$Total,
                                         p=HMD_Best_withRotatedLC_Female$p[HMD_Best_withRotatedLC_Female$Country=="ISL"],
                                         rotation_start_e0 = HMD_Best_withRotatedLC_Female$e0l[HMD_Best_withRotatedLC_Female$Country=="ISL"],
                                         Gender = "female")

colnames(est_mx_female) <- min(as.numeric(colnames(DataOutput$mx_female_train))):1999

temp <- ForecastRotatedLC(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                e0_female_proj = NULL, e0_male_proj = NULL,
                                HMD_Best_withRotatedLC_Female$p[HMD_Best_withRotatedLC_Female$Country=="ISL"],
                                HMD_Best_withRotatedLC_Female$e0l[HMD_Best_withRotatedLC_Female$Country=="ISL"],
                                2035-1999, IsLog = FALSE)$est_mx_female

colnames(temp) <- 2000:2035

est_mx_female <-cbind(est_mx_female, temp)

```

Get the estimated $m_x$ for the training set and forecasted years for each country in a loop.

```{r}
CountryList <- unique(HMD_Best_withRotatedLC_Female$Country)

ResultListMale <- vector(mode = "list", length = length(CountryList))
names(ResultListMale) <- CountryList

ResultListFemale <- vector(mode = "list", length = length(CountryList))
names(ResultListFemale) <- CountryList

for (index in 1:length(CountryList)) {
  DataOutput <- CleanUp_HMD(CountryList[index], 1950, 1999, 2011)
  
  print(CountryList[index])
  
  # Male
  
  tryCatch(
    expr = {
      est_mx_male <- EstimateTrainingRotatedLC(DataOutput$mx_male_train, DataOutput$mx_female_train, 1, DataOutput$e0data$Total,
                                         p=HMD_Best_withRotatedLC_Male$p[HMD_Best_withRotatedLC_Male$Country==CountryList[index]],
                                         rotation_start_e0 = HMD_Best_withRotatedLC_Male$e0l[HMD_Best_withRotatedLC_Male$Country==CountryList[index]],
                                         Gender = "male")
      colnames(est_mx_male) <- min(as.numeric(colnames(DataOutput$mx_male_train))):1999
      
      temp <- ForecastRotatedLC(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                e0_female_proj = NULL, e0_male_proj = NULL,
                                HMD_Best_withRotatedLC_Male$p[HMD_Best_withRotatedLC_Male$Country==CountryList[index]],
                                HMD_Best_withRotatedLC_Male$e0l[HMD_Best_withRotatedLC_Male$Country==CountryList[index]],
                                2035-1999, IsLog = FALSE)$est_mx_male
      colnames(temp) <- 2000:2035
      est_mx_male <-cbind(est_mx_male, temp)
      
      ResultListMale[[index]] <- est_mx_male
    },
    error = function(error_condition){ 
        ResultListMale[[index]] <- error_condition
    })
    
    # Female
  
  tryCatch(
    expr = {
      est_mx_female <- EstimateTrainingRotatedLC(DataOutput$mx_male_train, DataOutput$mx_female_train, 1, DataOutput$e0data$Total,
                                         p=HMD_Best_withRotatedLC_Female$p[HMD_Best_withRotatedLC_Female$Country==CountryList[index]],
                                         rotation_start_e0 = HMD_Best_withRotatedLC_Female$e0l[HMD_Best_withRotatedLC_Female$Country==CountryList[index]],
                                         Gender = "female")
      colnames(est_mx_female) <- min(as.numeric(colnames(DataOutput$mx_female_train))):1999
      temp <- ForecastRotatedLC(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                e0_female_proj = NULL, e0_male_proj = NULL,
                                HMD_Best_withRotatedLC_Female$p[HMD_Best_withRotatedLC_Female$Country==CountryList[index]],
                                HMD_Best_withRotatedLC_Female$e0l[HMD_Best_withRotatedLC_Female$Country==CountryList[index]],
                                2035-1999, IsLog = FALSE)$est_mx_female
      colnames(temp) <- 2000:2035
      est_mx_female <-cbind(est_mx_female, temp)
      
      ResultListFemale[[index]] <- est_mx_female
    },
    error = function(error_condition){ 
        ResultListFemale[[index]] <- error_condition
    })
  
}
```

Save results to *Rda*:

```{r}
save(ResultListMale,file="RotatedLC_Male_mx.Rda")

save(ResultListFemale,file="RotatedLC_Female_mx.Rda")
```

Save results to *xlsx*:

```{r}
library(xlsx)

# Male
wb <- createWorkbook()
sheetnames <- names(ResultListMale)
sheets <- lapply(sheetnames, createSheet, wb = wb)
void <- Map(addDataFrame, ResultListMale, sheets)
saveWorkbook(wb, file = "RotatedLC_Male_mx.xlsx")

wb <- NULL
gc()

# Female
wb <- createWorkbook()
sheetnames <- names(ResultListFemale)
sheets <- lapply(sheetnames, createSheet, wb = wb)
void <- Map(addDataFrame, ResultListFemale, sheets)
saveWorkbook(wb, file = "RotatedLC_Female_mx.xlsx")
```

## 4. Get the estimated $m_x$ from the best LC with GAM for each country

Load the detailed *LC with GAM* results:

```{r}
HMD_LCwithGAM_Full <- read_csv("HMD_LCwithGAM_Full_LogMSE.csv")
HMD_LCwithGAM_Full <- as.data.frame(HMD_LCwithGAM_Full)
str(HMD_LCwithGAM_Full)
```

Adjust column names to be compatible with HMD_BestResults:

```{r}
colnames(HMD_LCwithGAM_Full)[5:6] <- colnames(HMD_BestResults[4:5])
str(HMD_LCwithGAM_Full)
```

Let's get the $e0l$ and $p$ parameters based on the test $MSE$ in the `HMD_BestResults` dataframe:

```{r}
# Males
HMD_Best_LCwithGAM_Male <- HMD_BestResults[,c("Country", "mse_male_test_LCwithGAM")]
HMD_Best_LCwithGAM_Male <- merge(HMD_Best_LCwithGAM_Male,
                                     HMD_LCwithGAM_Full[,c("Country",
                                                           "mse_male_test_LCwithGAM",
                                                           "basis_form",
                                                           "k_base")],
                                     by=c("Country", "mse_male_test_LCwithGAM"))

#----------------------------------------------------------------------------------------------------

# Females
HMD_Best_LCwithGAM_Female <- HMD_BestResults[,c("Country", "mse_female_test_LCwithGAM")]
HMD_Best_LCwithGAM_Female <- merge(HMD_Best_LCwithGAM_Female,
                                     HMD_LCwithGAM_Full[,c("Country",
                                                           "mse_female_test_LCwithGAM",
                                                           "basis_form",
                                                           "k_base")],
                                     by=c("Country", "mse_female_test_LCwithGAM"))

# Getting the simpler model in case of multiple optima

HMD_Best_LCwithGAM_Female <- HMD_Best_LCwithGAM_Female[-c(7,26),]

```

Sourcing the necessary functions.

```{r}
source("RotatedLC_MortRateExportFunctions.R")
```

Get the estimated $m_x$ for the training set and forecasted years for one country and one gender.

```{r}
country <- "CAN"

DataOutput <- CleanUp_HMD(country, 1950, 1990, 1999)



est_mx_female <- exp(EstimateTrainingwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = HMD_Best_LCwithGAM_Female$k_base[HMD_Best_LCwithGAM_Female$Country==country],
                                 basis_form = HMD_Best_LCwithGAM_Female$basis_form[HMD_Best_LCwithGAM_Female$Country==country],
                                 IsLog = TRUE)$est_mx_female)

colnames(est_mx_female) <- min(as.numeric(colnames(DataOutput$mx_female_train))):1990

temp <- exp(ForecastLCwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = HMD_Best_LCwithGAM_Female$k_base[HMD_Best_LCwithGAM_Female$Country==country],
                                 basis_form = HMD_Best_LCwithGAM_Female$basis_form[HMD_Best_LCwithGAM_Female$Country==country],
                                 ForecastPeriods = 2035-1990,
                                 IsLog = TRUE)$est_mx_female)

colnames(temp) <- 1991:2035

est_mx_female <-cbind(est_mx_female, temp)


est_mx_male <- exp(EstimateTrainingwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = HMD_Best_LCwithGAM_Male$k_base[HMD_Best_LCwithGAM_Male$Country==country],
                                 basis_form = HMD_Best_LCwithGAM_Male$basis_form[HMD_Best_LCwithGAM_Male$Country==country],
                                 IsLog = TRUE)$est_mx_male)

colnames(est_mx_male) <- min(as.numeric(colnames(DataOutput$mx_male_train))):1990

temp <- exp(ForecastLCwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = HMD_Best_LCwithGAM_Male$k_base[HMD_Best_LCwithGAM_Male$Country==country],
                                 basis_form = HMD_Best_LCwithGAM_Male$basis_form[HMD_Best_LCwithGAM_Male$Country==country],
                                 ForecastPeriods = 2035-1990,
                                 IsLog = TRUE)$est_mx_male)

colnames(temp) <- 1991:2035

est_mx_male <-cbind(est_mx_male, temp)



```

Get the estimated $m_x$ for the training set and forecasted years for each country in a loop.

```{r}
CountryList <- unique(HMD_Best_LCwithGAM_Female$Country)

ResultListMale <- vector(mode = "list", length = length(CountryList))
names(ResultListMale) <- CountryList

ResultListFemale <- vector(mode = "list", length = length(CountryList))
names(ResultListFemale) <- CountryList

for (index in 1:length(CountryList)) {
  DataOutput <- CleanUp_HMD(CountryList[index], 1950, 1990, 1999)
  
  print(CountryList[index])
  
  # Male
  
  tryCatch(
    expr = {
      est_mx_male <- exp(EstimateTrainingwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = HMD_Best_LCwithGAM_Male$k_base[HMD_Best_LCwithGAM_Male$Country==CountryList[index]],
                                 basis_form = HMD_Best_LCwithGAM_Male$basis_form[HMD_Best_LCwithGAM_Male$Country==CountryList[index]],
                                 IsLog = TRUE)$est_mx_male)
      
      colnames(est_mx_male) <- min(as.numeric(colnames(DataOutput$mx_male_train))):1990
      
      temp <- exp(ForecastLCwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = HMD_Best_LCwithGAM_Male$k_base[HMD_Best_LCwithGAM_Male$Country==CountryList[index]],
                                 basis_form = HMD_Best_LCwithGAM_Male$basis_form[HMD_Best_LCwithGAM_Male$Country==CountryList[index]],
                                 ForecastPeriods = 2035-1990,
                                 IsLog = TRUE)$est_mx_male)
      
      colnames(temp) <- 1991:2035
      
      est_mx_male <-cbind(est_mx_male, temp)
      
      ResultListMale[[index]] <- est_mx_male
    },
    error = function(error_condition){ 
        ResultListMale[[index]] <- error_condition
    })
    
    # Female
  
  tryCatch(
    expr = {
      est_mx_female <- exp(EstimateTrainingwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = HMD_Best_LCwithGAM_Female$k_base[HMD_Best_LCwithGAM_Female$Country==CountryList[index]],
                                 basis_form = HMD_Best_LCwithGAM_Female$basis_form[HMD_Best_LCwithGAM_Female$Country==CountryList[index]],
                                 IsLog = TRUE)$est_mx_female)
      colnames(est_mx_female) <- min(as.numeric(colnames(DataOutput$mx_female_train))):1990
      
      temp <- exp(ForecastLCwithGAM(mx_male = DataOutput$mx_male_train, mx_female = DataOutput$mx_female_train, nx=1,
                                 k_base = HMD_Best_LCwithGAM_Female$k_base[HMD_Best_LCwithGAM_Female$Country==CountryList[index]],
                                 basis_form = HMD_Best_LCwithGAM_Female$basis_form[HMD_Best_LCwithGAM_Female$Country==CountryList[index]],
                                 ForecastPeriods = 2035-1990,
                                 IsLog = TRUE)$est_mx_female)
      colnames(temp) <- 1991:2035
      
      est_mx_female <-cbind(est_mx_female, temp)
      
      ResultListFemale[[index]] <- est_mx_female
    },
    error = function(error_condition){ 
        ResultListFemale[[index]] <- error_condition
    })
  
}
```

Save results to *Rda*:

```{r}
save(ResultListMale,file="LCwithGAM_Male_mx.Rda")

save(ResultListFemale,file="LCwithGAM_Female_mx.Rda")
```

Save results to *xlsx*:

```{r}
library(xlsx)

# Male
wb <- createWorkbook()
sheetnames <- names(ResultListMale)
sheets <- lapply(sheetnames, createSheet, wb = wb)
void <- Map(addDataFrame, ResultListMale, sheets)
saveWorkbook(wb, file = "LCwithGAM_Male_mx.xlsx")

wb <- NULL
gc()

# Female
wb <- createWorkbook()
sheetnames <- names(ResultListFemale)
sheets <- lapply(sheetnames, createSheet, wb = wb)
void <- Map(addDataFrame, ResultListFemale, sheets)
saveWorkbook(wb, file = "LCwithGAM_Female_mx.xlsx")
```